---
title: "Tidy data"
output: github_document
---

```{r setup, echo = FALSE, message = FALSE}
library(tidyverse)
```

## PULSE data

```{r}
pulse_df =
  haven::read_sas("data/public_pulse_data.sas7bdat") |>
  janitor::clean_names() |>
  pivot_longer( #making a long format
    bdi_score_bl:bdi_score_12m, # the range of columns you want to pivot
    names_to = "visit", # change the columns' names to the values for a variable "visit"
    values_to = "bdi_score", # change the columns' values to the values of variable "bdi_score"
    names_prefix = "bdi_score_" # remove prename "bdi_score_" from all values
  ) |>
  mutate(
    visit = replace(visit, visit == "bl", "00m")
    # find the "visit" values that are "bl" and replace them with "00m"
  )

```

# Learning Assessment 
In the litters data, the variables `gd0_weight` and `gd18_weight` give the weight of the mother mouse on gestational days 0 and 18. Write a data cleaning chain that retains only `litter_number` and these columns; produces new variables `gd` and `weight`; and makes `gd` a numeric variable taking values `0` and `18` (for the last part, you might want to use `recode` …). Is this version “tidy”?

```{r}
litters_df =
  read_csv("data/FAS_litters.csv") |> 
  janitor::clean_names() |>
  select(litter_number, gd0_weight, gd18_weight) |>
  pivot_longer(
    gd0_weight:gd18_weight,
    names_to = "gd",
    values_to = "weight",
    names_prefix = "gd"
  ) |>
  mutate(
    gd = case_match(
      gd,
      "0_weight" ~ 0, # anytime see a value "0_weight", replace it w/ 0
      "18_weight" ~ 18,
    )
    # gd = replace(gd, gd == "0_weight", 0)
  )
  
```


## LoTR

Import LoTR words data

```{r}
fellowship_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "B3:D6") |>
  # "range" indicates the range of EXCEL you want to work on
  mutate(movie = "fellowship")

two_towers_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "F3:H6") |>
  mutate(movie = "two towers")

return_of_the_king_df = 
  readxl::read_excel("data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "return of the king")

lotr_df =
  bind_rows(fellowship_df, two_towers_df, return_of_the_king_df) |>
  janitor::clean_names() |>
  pivot_longer(
    male:female,
    names_to = "gender",
    values_to = "word"
  ) |>
  relocate(movie) # move the "movie" column to the first column

```

## Revisit FAS

```{r}
litters_df = 
  read_csv("data/FAS_litters.csv") |>
  janitor::clean_names() |>
  mutate(wt_gain = gd18_weight - gd0_weight) |>
  select(litter_number, group, wt_gain) |>
  separate(group, into = c("dose", "day_of_tx"), 3)
  # separate the column "group" into two columns "dose" and "day_of_tx"
  
pups_df =
  read_csv("data/FAS_pups.csv") |>
  janitor::clean_names() |>
  mutate(
    sex = case_match(
      sex,
      1 ~ "male",
      2 ~ "female"
    )
  )
  
fas_df =
  left_join(pups_df, litters_df, by = "litter_number")
# merge "pups_df" and "litters_df" into a new df "fas_df"
# by having same values in "litter_number", a column that both df has
# left_join() keeps all observations in "pups_df"
```

